---
title: "Religions in Africa"
author: "Yabra Muvdi"
date: "August 20, 2018"
output: html
---

```{r Setup, warning=FALSE, message=FALSE}
encoding: UTF-8
library(sf)
library(ggplot2)
library(tidyverse)
library(xlsx)
library(viridis)
library(scales)
library(ggrepel)
setwd("C:/Users/Yabra/Desktop/Instiglio Maps")

```

I start be loading and storing the Geographical Data Files (GDF) for the whole world.

```{r Geo Data File}
gdf <- st_read("gadm36_levels_shp/gadm36_0.shp")
glimpse(gdf)
```

Lets plot the world
```{r Whole world}
plot(gdf["NAME_0"])
```

Now I need to load interesting data to match with the map of the world. I am going to load a simple data set with information on religious practices in each African country.

```{r Religion data}
rel_africa <- read.xlsx("C:/Users/Yabra/Desktop/Instiglio Maps/religions.xlsx", sheetIndex = 1,
                        encoding = "UTF-8") #Encoding option used to handle special characters

glimpse(rel_africa)
```

Time to join geographical data with the information on religious practices that I just loaded.

```{r Join data, warning=FALSE}
#Make sure names are stored as characters and not factors
gdf$NAME_0 <- as.character(gdf$NAME_0)
rel_africa$NAME_0 <- as.character(rel_africa$NAME_0)

#Join both datasets
rel_africa_map <- gdf %>% left_join(rel_africa, by = "NAME_0")

#Get rid of data outside of Africa 
rel_africa_map <- rel_africa_map %>% filter(!is.na(Islam))

#Check if all data matched correctly
dim(rel_africa) # 56 rows
dim(rel_africa_map) # 53 rows

#Some countries in the religion database were not matched with their geo data
missing <- anti_join(rel_africa, rel_africa_map)
missing
```

Lets see what are the names of those countries that didnt match
```{r, results= FALSE}
gdf$NAME_0
```

Rename countries with different names

```{r Rename countries}
rel_africa$NAME_0[rel_africa$NAME_0 == "Republic of the Congo"] <- "Republic of Congo"
rel_africa$NAME_0[rel_africa$NAME_0 == "Réunion"] <- "Reunion"
rel_africa$NAME_0[rel_africa$NAME_0 == "The Gambia"] <- "Gambia"

#Join the data again
rel_africa_map <- gdf %>% left_join(rel_africa, by = "NAME_0")
rel_africa_map <- rel_africa_map %>% filter(!is.na(Islam))
```


```{r Plot religion}
plot(rel_africa_map["Islam"])
```

## Making beautiful maps

I will start be defining a basic Instiglio template for maps

```{r Instilgio maps}
theme_instiglio_maps <- theme_void() + 
  theme(panel.grid = element_line(color = "transparent"), 
        plot.title = element_text(color = "#404040",face = "bold", hjust=0), 
        plot.subtitle = element_text(color = "#404040", face = "italic", hjust=0),
        plot.caption = element_text(color = "#404040", size = 8))
```

Now I will create a new plot

```{r Beatuiful map}
religions_africa <- ggplot() +
  geom_sf(data = rel_africa_map, aes(color = Islam ,fill = Islam)) +
  theme_void() +
  labs(
    title = "Religions in Africa",
    subtitle = "",
    caption = "Wikipedia",
    color = "Islam",
    fill = "Islam") + 
  theme_instiglio_maps +
    scale_fill_gradient2(high = "#c0504d", low = "#4f81bd" , midpoint = 50) + 
    scale_color_gradient2(high = "#c0504d", low = "#4f81bd", midpoint = 50)

religions_africa
```

However, I am assuming that the population that is Muslim, is Christian

## A different approach to the map

I am going to create a new variable that contains the religion practiced by the majority of the population in each country

```{r Recoding data}
rel_africa_map <- rel_africa_map %>% 
  mutate(religion = if_else(Christianity > 50, "Christianity", ""))

rel_africa_map <- rel_africa_map %>% 
  mutate(religion = if_else(Islam > 50, "Islam", religion))


rel_africa_map$Traditional.religions <- as.numeric(rel_africa_map$Traditional.religions)

rel_africa_map <- rel_africa_map %>% 
  mutate(religion = if_else(Traditional.religions > 50, "Traditional", religion))

rel_africa_map <- rel_africa_map %>% 
  mutate(religion = if_else(religion == "", "No majority", religion))

```



```{r Final map}
colors <- c("#4f81bd","#c0504d", "#1f6862", "#c4bd97")

religions_africa2 <- ggplot() +
  geom_sf(data = rel_africa_map, aes(color = religion, fill = religion)) +
  theme_void() + theme_instiglio_maps +
  labs(
    title = "Majority Religions in African Countries",
    subtitle = "",
    caption = "Wikipedia, 2018",
    color = "",
    fill = "") + 
    scale_color_manual(values = alpha(colors, 0.7), aesthetics = c("colour", "fill"))

religions_africa2
```